Robust Control of a Fluxonium Qubit

### 2020-11-28 paper/5static.tex
solid progress through sec4

\section{Robustness to Static Parameter Deviations \label{sec:static}}
We have formulated the quantum optimal control
problem as an open loop optimization problem, i.e.
feedback from the experiment is not incorporated in optimization.
However, the device typically deviates from the Hamiltonian we use in optimization,
leading to poor experimental performance. We combat errors
of this form using robust control techniques,
making the state evolution insensitive
to Hamiltonian parameter deviations. As an example
we mitigate errors arising from the drift and finite measurement
precision of the qubit frequency $\tilde{f_{q}} = f_{q} \pm \sigma_{f_{q}}$ \todo{maybe 
reference \eqref{eq:hamiltonian} here?}.
We consider three robust control techniques.
The first is the sampling method, which has
been proposed previously in the context of QOC
\cite{rembold2020introduction, reinhold2019controlling, carvalho2020error}. In the
sampling method, multiple states are evolved under distinct deviant dynamics
to capture the effect of parameter deviations. We also
study the unscented sampling method, which uses the unscented
transformation to accurately propagate a distribution
representing the uncertainty in an evolving state
due to a parameter deviation.
The unscented transformation was designed for nonlinear Kalman
filtering and is frequently utilized in robust control
\cite{howell2020direct, julier2004unscented, lee2013sigma, manchester2016derivative}.
Finally, we propose the derivative method. Derivative
information encoding the sensitivity of the
state trajectory with respect to the deviant parameter
is used to penalize state trajectory deviations.

\todo{I think you can make this more concrete and simpler to understand by just defining
the new augmented state vector and explaining you stack multiple copies of the state vector
together, each evolving with a different static parameter. In this case there's only one 
parameter being changed, so I'd just state it explicitly.}
In the sampling method, sample states evolve under a
Hamiltonian where a parameter is replaced by
a deviant value. We propagate the additional states $\ket{\psi^{\pm}}$ with
deviant values $\lambda^{\pm} = \lambda \pm \sigma_{\lambda}$.
The gate error of each sample state is 
penalized at each knot point
$1 - {\lvert \braket{\psi^{\pm}_{k} | \psi_{f}} \rvert}^{2}$.
Each initial state is used to initialize two
sample states $\ket{\psi^{\pm}}$.
The initial states are chosen so
that their outer products span the operators on the
Hilbert space $\{\ket{0}, \ket{1}, (\ket{0} + i\ket{1}) / \sqrt{2},
(\ket{0} - \ket{1}) / \sqrt{2}\}$ \cite{chow2009randomized}.
For this method the number of states in the augmented state vector
scales as $O(dn^{2})$ \todo{is this right?}
because each of $n^{2}$ initial
states is represented by $2d$ samples. $n$ is both the dimension of the Hilbert space
and the size of the state vector. $d$ is the number of deviant parameters.

@@

The strategy of the unscented sampling method is to
propagate an ensemble of sample states (sigma points) which represent a distribution
over every element of the initial state. The distribution
models the uncertainty in each state element arising from the
parameter deviation. The ensemble consists of $4n + 2d$ states.
Each state in the
ensemble is propagated to the next
knot point using separate deviant dynamics. Then, the mean and covariance of
the propagated ensemble is calculated. New states are sampled from the distribution
given by the calculated statistics for propagation to the next knot point.
This resampling procedure, the unscented transformation, accurately propagates
first and second moments of the distribution and ensures the points
lie on the $\sqrt{2n}$\textsuperscript{th}
covariance contour at each knot point. A detailed update procedure is given
in Appendix \ref{appendix:unscented}. Each initial state from the operator basis
is represented with a distribution of $4n + 2d$ samples. Hence,
the number of states in the augmented state vector scales as $O(n^{3} + dn^{2})$.
The gate error of each sample state is penalized as in the sampling method.
\todo{I think this section needs some more detailed explanation.}

The derivative method draws on the intuition that
the sensitivity of the state evolution to a parameter
$\lambda$ is encoded in the $l$\textsuperscript{th}-order
derivative of the state with respect to that parameter
$\ket{\partial_{\lambda}^{l} \psi}$. The $m$\textsuperscript{th}-order
derivative method minimizes the norm of the first $m$
state derivatives with a quadratic cost at each knot point
${\lvert \braket{\partial^{l}_{\lambda} \psi | \partial^{l}_{\lambda} \psi}
  \rvert}^{2}$, $l \in \{1, \dots, m\}$.
The state derivatives could be obtained with backwards mode differentiation.
Naive automatic differentiation would compute
the state derivative at all $1, \dots, k - 1$ knot points
to obtain the state derivative at knot point $k$.
For a single state derivative and $N$ knot points this
requires $O(N^2)$ matrix multiplications.
Instead, we forward propagate the state derivatives in the
augmented state vector under coupled dynamics, resulting in
$O(N)$ matrix multiplications. For example, the dynamics
for the 1\textsuperscript{st}-order derivative method are
\begin{align}
  i \hbar \frac{d}{dt} \ket{\psi} &= H \ket{\psi}\\
  i \hbar \frac{d}{dt} \ket{\partial_{\lambda}\psi} &=
  H \ket{\partial_{\lambda} \psi} +
  (\partial_{\lambda} H) \ket{\psi}
\end{align}
\todo{inconsistent use of $\hbar$ vs $h$.}
Exponential integrators that account for the non-linear
term may be used to efficiently integrate the coupled dynamics,
see Appendix \ref{appendix:derivative}.
For this method $m$ state derivatives are associated with each
initial state from the operator basis.
So, the number of states in the augmented state vector scales as
$O(dmn^{2})$. Additionally, we note that the number of initial
states and samples we have employed for each of the three
methods follows standard perscriptions. However, knowledge
of the problem can typically be used to reduced the number
of required samples. \todo{I think this also needs some more explanation. It's not 
clear exactly what you're doing. You're penalizing the derivatives of the dynamics with
respect to the uncertain parameter in the objective, so why do you need to augment the 
state vector? Or are you adding the derivatives to the state vector so that you have 
standard quadratic penalty terms in the objective? Again, I think this confusion would 
be easily cleared up by explicitly writing down your new dynamics, objective function, 
and any additional constraints you impose on the system. Since this is really the meat of 
the paper, it's worth spending the time to make sure this is clear and easy to understand,
especially for those who aren't very familiar with trajectory optimization or optimization 
in general. Also, I think you should explicitly state the size of the problem you're
actually solving. Might be worth putting in a table, along with some timing results, to 
give an idea of how well these methods scale.}

@@

Additionally, there are multiple analytic methods 
to mitigate errors due to parameter deviations
including composite pulse sequences \cite{merrill2014progress},
the DRAG scheme \cite{krantz2019quantum}, and
geometric phase considerations
\cite{xu2020nonadiabatic, han2020experimental}.
Composite pulse sequences are derived by computing
the erroneous rotation arising from the parameter deviation
with a finite order magnus expansion.
Pulses are then suitably composed to eliminate the
error. In principle the error may be eliminated
to arbitrary order with sufficiently many pulses \cite{merill2014progress},
which mirrors the negative correlation between
gate error and gate duration we observe with the derivative method.
It is difficult to choose an appropriate composite pulse
for the problem studied here,
so we propose comparisons between composite pulses and numerical techniques
for future work.

>>>>

\section{Robustness to Static Parameter Deviations \label{sec:static}}
We have formulated the quantum optimal control
problem as an open loop optimization problem, i.e.
feedback from the experiment is not incorporated in optimization.
However, the device typically deviates from the Hamiltonian we use in optimization,
leading to poor experimental performance. We combat errors
of this form using robust control techniques,
making the state evolution insensitive
to Hamiltonian parameter deviations. As an example,
we mitigate errors arising from the drift and finite measurement
precision of the qubit frequency in \eqref{eq:hamiltonian}.
We consider three robust control techniques:
the sampling method, the unscented sampling method,
and the derivative method.

The sampling method optimizes over copies of a quantum state,
called samples.
Each sample is associated with a distinct deviant parameter value.
Two samples $\psi_{\pm}$
are appended to the augmented state vector \eqref{eq:astatecontrols}
for each of four initial states
$\{\ket{0}, \ket{1}, (\ket{0} + i\ket{1}) / \sqrt{2},
(\ket{0} - \ket{1}) / \sqrt{2}\}$ which are chosen
so that their outer products span the operators on the
Hilbert space  \cite{chow2009randomized}.
The discrete dynamics \eqref{eq:dyn_con} is modified so that the
samples evolve under the Hamiltonian given in \eqref{eq:hamiltonian}
with $f_{q} \gets f_{q} \pm \sigma_{f_{q}}$. The final state
for each sample is the image of the initial state
under the desired gate, and deviations from the final state are
penalized in the objective \eqref{eq:costfun}.

@@

For the unscented sampling method, $2(2n + d)$ samples
are appended to the augmented state vector \eqref{eq:astatecontrols}
for each initial state from the operator basis. Here $2n = 4$ is the
dimension of the Hilbert space under the isomorphism \eqref{eq:isomorphism}
and $d = 1$ is the number of deviant parameters. The sample
states are used to represent a Gaussian distribution over all $2n$
elements of the initial state, modeling
the uncertainty in the state as a result of the uncertainty in
the deviant parameter. The unscented transformation
is applied to the sample ensemble at each knot point in the discrete dynamics
\eqref{eq:dyn_con} to ensure the first and second moments of the distribution
are accurately propagated. Each sample evolves under the Hamiltonian given in
\eqref{eq:hamiltonian}
where the qubit frequency is replaced by a deviant value determined by the
statistics of the sample ensemble. As in the sampling method,
deviations from the final state are penalized according to \eqref{eq:costfun}.
A detailed update procedure for this method is given
in Appendix \ref{appendix:unscented}.

The derivative method draws on the intuition that
the sensitivity of the state evolution to the deviant parameter
is encoded in the ($l$\textsuperscript{th}-order)
derivative of the state with respect to that parameter
$\partial_{f_{q}}^{l} \psi$.

The $m$\textsuperscript{th}-order
derivative method minimizes the norm of the first $m$
state derivatives with a quadratic cost at each knot point
${\lvert \braket{\partial^{l}_{\lambda} \psi | \partial^{l}_{\lambda} \psi}
  \rvert}^{2}$, $l \in \{1, \dots, m\}$.
The state derivatives could be obtained with backwards mode differentiation.
Naive automatic differentiation would compute
the state derivative at all $1, \dots, k - 1$ knot points
to obtain the state derivative at knot point $k$.
For a single state derivative and $N$ knot points this
requires $O(N^2)$ matrix multiplications.
Instead, we forward propagate the state derivatives in the
augmented state vector under coupled dynamics, resulting in
$O(N)$ matrix multiplications. For example, the dynamics
for the 1\textsuperscript{st}-order derivative method are
\begin{align}
  i \hbar \frac{d}{dt} \ket{\psi} &= H \ket{\psi}\\
  i \hbar \frac{d}{dt} \ket{\partial_{\lambda}\psi} &=
  H \ket{\partial_{\lambda} \psi} +
  (\partial_{\lambda} H) \ket{\psi}
\end{align}
\todo{inconsistent use of $\hbar$ vs $h$.}
Exponential integrators that account for the non-linear
term may be used to efficiently integrate the coupled dynamics,
see Appendix \ref{appendix:derivative}.
\todo{I think this also needs some more explanation. It's not 
clear exactly what you're doing. You're penalizing the derivatives of the dynamics with
respect to the uncertain parameter in the objective, so why do you need to augment the 
state vector? Or are you adding the derivatives to the state vector so that you have 
standard quadratic penalty terms in the objective? Again, I think this confusion would 
be easily cleared up by explicitly writing down your new dynamics, objective function, 
and any additional constraints you impose on the system. Since this is really the meat of 
the paper, it's worth spending the time to make sure this is clear and easy to understand,
especially for those who aren't very familiar with trajectory optimization or optimization 
in general. Also, I think you should explicitly state the size of the problem you're
actually solving. Might be worth putting in a table, along with some timing results, to 
give an idea of how well these methods scale.}

@@

Additionally, there are multiple analytic methods 
to mitigate errors due to parameter deviations
including composite pulse sequences \cite{merrill2014progress},
the DRAG scheme \cite{krantz2019quantum}, and
geometric phase considerations
\cite{xu2020nonadiabatic, han2020experimental}.
Composite pulse sequences are derived by computing
the erroneous rotation arising from the parameter deviation
with a finite order magnus expansion.
Pulses are then suitably composed to eliminate the
error. In principle the error may be eliminated
to arbitrary order with sufficiently many pulses \cite{merrill2014progress},
which mirrors the negative correlation between
gate error and gate duration we observe with the derivative method.
It is difficult to choose an appropriate composite pulse
for the problem studied here,
so we propose comparisons between composite pulses and numerical techniques
for future work.
